From Freebase to Wikidata: The Great Migration

### 2016-01-29 paper.tex
Reworked Section 5.1.

``Useful'' facts are the triples that remain
after filtering out all triples about IDs (\texttt{/type/object/id}),
types \linebreak (\texttt{/type/object/type}), labels, descriptions \linebreak
(\texttt{/type/object/name},
\texttt{/common/topic/description}, \emph{etc.}), \linebreak and triples removed
for licensing reasons (the latter were about 42~million, see \autoref{sec:licensing}).
The triples removed in this step are not considered to be potentially useful for
Wikidata, since IDs, labels, and descriptions are not handled through statements.

As a~direct comparison, here are the raw statistics for Wikidata as of August 2015.

@@

We also cannot compare the number of facts and the number of statements directly,
even considering the lower number of topics.
Freebase encodes its knowledge far more redundantly than Wikidata.
To illustrate this, properties in Freebase often have a~\emph{reverse} property
that is used in order to be able to traverse the Freebase graph easily in both directions.
For example, the property \texttt{/people/person/place\_of\_birth} has the corresponding
reverse property \linebreak \texttt{/location/location/people\_born\_here}
that encodes exactly the same semantic information.
Such reverse properties sometimes exist in Wikidata
(like \texttt{children} for \texttt{father} and \texttt{mother}),
but are far less common and also not automatically handled.

Another point is that CVTs use a~lot of facts to convey the exact same information
that can be represented with a~single Wikidata statement.
For example, to encode that Barack Obama is the president
of the United States since January~20, 2009,
Freebase requires more than six facts as shown in \autoref{fig:cvt-obama}.
If we attempted to encode Wikidata statements as if they were Freebase facts, \emph{i.e.},
by removing sources, representing statements with qualifiers using CVTs,
and adding reverse properties, this would lead to a~number of 110~million facts,
\emph{i.e.} an increase of 167\% over the raw number of statements.
The representation of Wikidata in RDF and its impact on the number of triples
is discussed in more detail in~\cite{reification-wikidata}.

In a~considerable amount of cases, Freebase also contains duplicate data.
For example, many cities in the United States have duplicate CVTs for the population,
where the only difference is the encoding of the source.

>>>>

``Useful'' facts are the triples that remain
after filtering out all triples about labels
(\texttt{/type/object/name}, identifiers (\texttt{/type/object/id}),
types (\texttt{/type/object/type}), descript\-ions
(\texttt{/common/topic/description}), \emph{etc.}), and triples removed
for licensing reasons (the latter were about 42~million, see \autoref{sec:licensing}).
The triples removed in this step are not considered to be potentially useful for
Wikidata, since IDs, labels, and descriptions are not handled through statements.

As a~direct comparison, in the following we provide the raw statistics for Wikidata as of August 2015.

@@

We also cannot compare the number of facts and the number of statements directly,
even considering the lower number of topics.
Freebase encodes its knowledge far more redundantly than Wikidata.
To illustrate this, properties in Freebase often have a~\emph{reverse} property
that is used in order to be able to traverse the Freebase graph easily in both directions.
For example, the property \texttt{/people/person/place\-\_of\_birth} has the corresponding
reverse property \linebreak \texttt{/location/location/people\_born\_here}
that encodes exactly the same semantic information.
Such reverse properties sometimes exist in Wikidata
(like \texttt{children} for \texttt{father} and \texttt{mother}),
but are far less common and also not automatically handled.

Another point is that CVTs use a~lot of facts to convey the exact same information
that can be represented with a~single Wikidata statement.
For example, to encode that Barack Obama is the president
of the United States since January~20, 2009,
Freebase requires more than six facts as shown in \autoref{fig:cvt-obama}.
If we attempted to encode Wikidata statements as if they were Freebase facts, \emph{i.e.},
by removing sources, representing statements with qualifiers using CVTs,
and adding reverse properties, this would lead to a~number of 110~million facts,
\emph{i.e.} an increase of 167\% over the raw number of statements.
The representation of Wikidata in RDF and its impact on the number of triples
is discussed in more detail in~\cite{reification-wikidata}.
In a~considerable amount of cases, Freebase also contains duplicate data.
For example, many cities in the United States have duplicate CVTs for the population,
where the only difference is the encoding of the source.
